### 2014-12-14 text/AppendixB.tex
Finish tweaks in flat coeffects (Chapter 3)
Proofreading and English conventions (Chapter 4)
Proofreading and editing (Chapter 5)
Proving things (Chapter 5)

\chapter{Appendix A} 
\label{ch:appendix}

@@

@@

\paragraph{(var)} Given the following derivation using (\emph{var}):
\[
\inference
  { }
  {\coctx{\Gamma_1, y:\tau, \Gamma_2}{\cunit} \vdash y : \tau }
\]
There are two cases depending on whether $y$ is the variable $x$ or not:
\begin{itemize}
\item[--] If $y=x$ then also $\tau = \tau_s$ and thus $\subst{y}{x}{e_s} = e_s$. Using the assumption,
implicit weakening and the fact that $\cunit$ is a unit of $\cseq$:
\[
\inference
 {\inference 
  {\coctx{\Gamma}{\cclrd{S}} \vdash \subst{y}{x}{e_s} : \tau_s }
  {\coctx{\Gamma_1, \Gamma, \Gamma_2}{\cclrd{S}} \vdash \subst{y}{x}{e_s} : \tau } }
 {\coctx{\Gamma_1, \Gamma, \Gamma_2}{\cunit\,\cseq\,\cclrd{S}} \vdash \subst{y}{x}{e_s} : \tau } 
\]
\item[--] If $y\neq x$ then $\subst{y}{x}{e_s} = y$. Using the fact that $\cunit$ is the bottom element
and sub-coeffecting:
\[
\inference
  {\coctx{\Gamma_1, y:\tau, \Gamma_2}{\cunit} \vdash y : \tau }
  {\coctx{\Gamma_1, y:\tau, \Gamma_2}{\cunit\,\cseq\,\cclrd{S}} \vdash y : \tau }
\]
\end{itemize}

@@

\paragraph{(app)} Given the following typing derivation using (\emph{app}):
\[
\inference
  {\coctx{\Gamma_1,x:\tau_s,\Gamma_2}{\cclrd{r}} \vdash e_1 : \tau_1 \xrightarrow{\cclrd{t}} \tau_2 &
   \coctx{\Gamma_1,x:\tau_s,\Gamma_2}{\cclrd{s}} \vdash e_2 : \tau_1 }
  {\coctx{\Gamma_1,x:\tau_s,\Gamma_2}{\cclrd{r} \;\cpar\; (\cclrd{s} \,\cseq\, \cclrd{t})} \vdash e_1~e_2 : \tau_2}
\]
From the induction hypothesis, we have that:
\[
\begin{array}{l}
 \coctx{\Gamma_1, \Gamma, \Gamma_2}{\cclrd{r}\,\cseq\,\cclrd{S}} \vdash \subst{e_1}{x}{e_s} : \tau_1 \xrightarrow{\cclrd{t}} \tau_2 \\
 \coctx{\Gamma_1, \Gamma, \Gamma_2}{\cclrd{s}\,\cseq\,\cclrd{S}} \vdash \subst{e_2}{x}{e_s} : \tau_1
\end{array}\quad(*)
\]
Now using (\emph{app}) rule and the fact that $\cpar\,=\,\cseq$, associativity, commutativity and idempotence
(note that all three properties are needed):
\[
\inference
 { (*) }
 {\inference
   { \coctx{\Gamma_1,\Gamma,\Gamma_2}{(\cclrd{r}\,\cseq\,\cclrd{S}) \;\cpar\; ((\cclrd{s}\,\cseq\,\cclrd{S}) \,\cseq\, \cclrd{t})} 
       \vdash \subst{e_1}{x}{e_s}~\subst{e_2}{x}{e_s} : \tau_2}
   { \coctx{\Gamma_1,\Gamma,\Gamma_2}{(\cclrd{r} \;\cpar\; (\cclrd{s} \,\cseq\, \cclrd{t}))\;\cseq\;\cclrd{S}} 
       \vdash \subst{(e_1~e_2)}{x}{e_s} : \tau_2} }
\]

\paragraph{(let)} Given the following typing derivation using (\emph{let}):
\[
\inference
  { \coctx{\Gamma_1,x:\tau_s,\Gamma_2}{\cclrd{r}} \vdash e_1 : \tau_1 &
    \coctx{\Gamma_1,x:\tau_s,\Gamma_2, y:\tau_1}{\cclrd{s}} \vdash e_2 : \tau_2}
  {\coctx{\Gamma_1,x:\tau_s,\Gamma_2}{\cclrd{s} \;\cpar\; (\cclrd{s} \,\cseq\, \cclrd{r})} \vdash \kvd{let}~y=e_1~\kvd{in}~e_2 : \tau_2 }
\]
From the induction hypothesis, we have that:
\[
\begin{array}{l}
 \coctx{\Gamma_1, \Gamma, \Gamma_2}{\cclrd{r}\,\cseq\,\cclrd{S}} \vdash \subst{e_1}{x}{e_s} : \tau_1 \\
 \coctx{\Gamma_1, \Gamma, \Gamma_2, y:\tau_1}{\cclrd{s}\,\cseq\,\cclrd{S}} \vdash \subst{e_2}{x}{e_s} : \tau_2
\end{array}\quad(\dagger)
\]
Now using (\emph{let}) rule and similarly to the (\emph{app}) case:
\[
\inference
 { (\dagger) }
 { \inference
   { \coctx{\Gamma_1, \Gamma, \Gamma_2}{(\cclrd{s}\,\cseq\,\cclrd{S}) \;\cpar\; ((\cclrd{s}\,\cseq\,\cclrd{S}) \,\cseq\, (\cclrd{r}\,\cseq\,\cclrd{S}))} 
         \vdash \kvd{let}~y=\subst{e_1}{x}{e_s}~\kvd{in}~\subst{e_2}{x}{e_s} : \tau_2 } 
   { \coctx{\Gamma_1, \Gamma, \Gamma_2}{(\cclrd{s} \;\cpar\; (\cclrd{s} \,\cseq\, \cclrd{r})) \,\cseq\, \cclrd{S}} 
         \vdash \subst{(\kvd{let}~y=e_1~\kvd{in}~e_2)}{x}{e_s} : \tau_2 } }
\]
\end{proof}

%
% (e2 + (e1 * e2)) + (e3 * f)
% ----(glet)------
% ----------(app)------------
%
% = e2 + e1*e2 + e3*f
%  
% ~~~>
%
% (e2 + (e3 * f)) + ((e2 + (e3 * f)) * e1)
% ----(app)------    ----(app)-----
% ---------------(glet)-------------------
%
% = e2 + e3*f + e2*f + e1*e3*f

% e3 + ((e2 + (e1 * e2)) * e3)
%       ------(glet)---
% ---------(glet)-------------
%
% = e3 + e2*e3 + e1*e2*e3
%  
% ~~~>
%
% (e2 + (e3 * e2)) + (e1 * (e2 + (e3 * e2)))
% -----(glet)-----         -----(glet)-----
% ------------------(glet)------------------
%
% = e2 + e3*e2 + e1*e2 + e1*e2*e3

% Data-flow: e3 + (e2 + e1) = (e2 + e3) + e1
% Liveness:  e3 = e3
% Resources: OK

%---------------------------------------------------------------------------------------------------

% \section{Internalized substitution}

% \subsection{First transformation}

% \begin{equation*}
% (\kvd{glet}~x=e_1~\kvd{in}~e_2)~e_3 \leadsto \kvd{glet}~x=e_1~\kvd{in}~(e_2~e_3)
% \end{equation*}

% \begin{equation*}
% \tyrule{app}
%   { \tyrule{glet}
%       { \coctx{\Gamma}{\cclrd{s}} \vdash e_1 : \tau_1 &
%         \coctx{\Gamma, x\!:\!\tau_1}{\cclrd{r}} \vdash e_2 : \tau_3 \xrightarrow{\cclrd{t}} \tau_2 }
%       { \coctx{\Gamma}{\cclrd{r}\;\cpar\;({\cclrd{s}\,\cseq\,\cclrd{r})}} \vdash \kvd{glet}~x=e_1~\kvd{in}~e_2 : \tau_3 \xrightarrow{\cclrd{t}} \tau_2 } &
%     \coctx{\Gamma}{\cclrd{u}} \vdash e_3 : \tau_3  }
%   { \coctx{\Gamma}{(\cclrd{r}\;\cpar\;(\cclrd{s}\,\cseq\,\cclrd{r})) \;\cpar\; (\cclrd{u}\,\cseq\,\cclrd{t}) } \vdash (\kvd{glet}~x=e_1~\kvd{in}~e_2)~e_3 : \tau_2 }
% \end{equation*}

% to

% \begin{equation*}
% \tyrule{glet}
%   { \coctx{\Gamma}{\cclrd{s}} \vdash e_1 : \tau_1 &
%     \tyrule{app}
%       { \coctx{\Gamma, x\!:\!\tau_1}{\cclrd{r}} \vdash e_2 : \tau_3 \xrightarrow{\cclrd{t}} \tau_2 & 
%         \coctx{\Gamma}{\cclrd{u}} \vdash e_3 : \tau_3 }
%       { \coctx{\Gamma}{\cclrd{r}\;\cpar\;(\cclrd{u}\,\cseq\,\cclrd{t})} \vdash e_2~e_3 : \tau_2 } }
%   { \coctx{\Gamma}{ (\cclrd{r}\;\cpar\;(\cclrd{u}\,\cseq\,\cclrd{t})) \;\cpar\; (\cclrd{s}\,\cseq\,(\cclrd{r}\;\cpar\;(\cclrd{u}\,\cseq\,\cclrd{t})))  }
%       \vdash \kvd{glet}~x=e_1~\kvd{in}~(e_2~e_3) : \tau_2 }
% \end{equation*}

% meaning
% \begin{equation*}
% \begin{array}{l}
% (\cclrd{r}\;\cpar\;(\cclrd{s}\,\cseq\,\cclrd{r})) \;\cpar\; (\cclrd{u}\,\cseq\,\cclrd{t}) =
% \end{array}
% \end{equation*}

% \subsection{Second transformation}

% Second transformation

% \begin{equation*}
% \tyrule{glet}
%   { \coctx{\Gamma}{\cclrd{s}} \vdash e_s : \tau_s &
%     \coctx{\Gamma, x\!:\!\tau_1}{\cclrd{r}} \vdash e_r : \tau_r &
%     \coctx{\Gamma, x\!:\!\tau_1}{\cclrd{t}} \vdash e_t : \tau_t }
%   { \coctx{\Gamma}{
%       \cclrd{t}
%       \;\cpar\;
%       (  (\cclrd{r}\;\cpar\;(\cclrd{s}\,\cseq\,\cclrd{r}))
%          \,\cseq\, 
%          \cclrd{t} )
%     } 
%     \vdash \kvd{glet}~x_r=(\kvd{glet}~x_s=e_s~\kvd{in}~e_r)~\kvd{in}~e_t : \tau_t }
% \end{equation*}

% or

% \begin{equation*}
% \tyrule{glet}
%   { \coctx{\Gamma}{\cclrd{s}} \vdash e_s : \tau_s &
%     \coctx{\Gamma, x\!:\!\tau_1}{\cclrd{r}} \vdash e_r : \tau_r &
%     \coctx{\Gamma, x\!:\!\tau_1}{\cclrd{t}} \vdash e_t : \tau_t }
%   { \coctx{\Gamma}{
%       (\cclrd{t}\;\cpar\;(\cclrd{r}\,\cseq\,\cclrd{t})) 
%       \;\cpar\;
%       (  \cclrd{s}
%          \,\cseq\, 
%          (\cclrd{t}\;\cpar\;(\cclrd{r}\,\cseq\,\cclrd{t})) )
%     } 
%     \vdash \kvd{glet}~x_s=e_s~\kvd{in}~(\kvd{glet}~x_r=e_r~\kvd{in}~e_t) : \tau_t }
% \end{equation*}

% \begin{equation*}
% \begin{array}{l}
%   \cclrd{t}
%   \;\cpar\;
%   (  (\cclrd{r}\;\cpar\;(\cclrd{s}\,\cseq\,\cclrd{r}))
%      \,\cseq\, 
%      \cclrd{t} ) = 
% \\
%   \cclrd{t}
%   \;\cpar\;
%   (\cclrd{r}\,\cseq\, \cclrd{t})
%   \;\cpar\;
%   (\cclrd{s}\,\cseq\,\cclrd{r}\,\cseq\, \cclrd{t}) = 
% \\
%   \cclrd{s}\,\cseq\,\cclrd{r}\,\cseq\, \cclrd{t} 
% \end{array}
% \end{equation*}

% \begin{equation*}
% \begin{array}{l}
%   (\cclrd{t}\;\cpar\;(\cclrd{r}\,\cseq\,\cclrd{t})) 
%   \;\cpar\;
%   (  \cclrd{s}
%      \,\cseq\, 
%      (\cclrd{t}\;\cpar\;(\cclrd{r}\,\cseq\,\cclrd{t})) ) = 
% \\     
%   \cclrd{t}\;\cpar\;(\cclrd{r}\,\cseq\,\cclrd{t})
%   \;\cpar\;
%   (\cclrd{s} \,\cseq\, \cclrd{t})
%   \;\cpar\;
%   (\cclrd{s} \,\cseq\, \cclrd{r}\,\cseq\,\cclrd{t})  = 
% \\     
%   \cclrd{s} \,\cseq\, \cclrd{r}\,\cseq\,\cclrd{t}
% \end{array}
% \end{equation*}

% require

% \begin{equation*}
% \cclrd{r}\;\cpar\;(\cclrd{r} \,\cseq\, \cclrd{s}) = \cclrd{r} \,\cseq\, \cclrd{s}
% \end{equation*}

>>>>

\chapter{Appendix B} 
\label{ch:appendix}

@@

% =================================================================================================
%
%      #####   ##            #    
%      #        #            #    
%      #        #     ###   ####  
%      ####     #        #   #    
%      #        #     ####   #    
%      #        #    #   #   #  # 
%      #       ###    ####    ##  
%                               
% =================================================================================================

@@

\paragraph{(var)} Given the following derivation using (\emph{var}):
\[
\inference
  { }
  {\coctx{\Gamma_1, y:\tau, \Gamma_2}{\cunit} \vdash y : \tau }
\]
There are two cases depending on whether $y$ is the variable $x$ or not:
\begin{itemize}
\item If $y=x$ then also $\tau = \tau_s$ and thus $\subst{y}{x}{e_s} = e_s$. Using the assumption,
implicit weakening and the fact that $\cunit$ is a unit of $\cseq$:
\[
\inference
 {\inference 
  {\coctx{\Gamma}{\cclrd{S}} \vdash \subst{y}{x}{e_s} : \tau_s }
  {\coctx{\Gamma_1, \Gamma, \Gamma_2}{\cclrd{S}} \vdash \subst{y}{x}{e_s} : \tau } }
 {\coctx{\Gamma_1, \Gamma, \Gamma_2}{\cunit\,\cseq\,\cclrd{S}} \vdash \subst{y}{x}{e_s} : \tau } 
\]
\item If $y\neq x$ then $\subst{y}{x}{e_s} = y$. Using the fact that $\cunit$ is the bottom element
and sub-coeffecting:
\[
\inference
  {\coctx{\Gamma_1, y:\tau, \Gamma_2}{\cunit} \vdash y : \tau }
  {\coctx{\Gamma_1, y:\tau, \Gamma_2}{\cunit\,\cseq\,\cclrd{S}} \vdash y : \tau }
\]
\end{itemize}

@@

\paragraph{(app)} Given the following typing derivation using (\emph{app}):
\[
\inference
  {\coctx{\Gamma_1,x:\tau_s,\Gamma_2}{\cclrd{r}} \vdash e_1 : \tau_1 \xrightarrow{\cclrd{t}} \tau_2 &
   \coctx{\Gamma_1,x:\tau_s,\Gamma_2}{\cclrd{s}} \vdash e_2 : \tau_1 }
  {\coctx{\Gamma_1,x:\tau_s,\Gamma_2}{\cclrd{r} \;\cpar\; (\cclrd{s} \,\cseq\, \cclrd{t})} \vdash e_1~e_2 : \tau_2}
\]
From the induction hypothesis, we have that:
\[
\begin{array}{l}
 \coctx{\Gamma_1, \Gamma, \Gamma_2}{\cclrd{r}\,\cseq\,\cclrd{S}} \vdash \subst{e_1}{x}{e_s} : \tau_1 \xrightarrow{\cclrd{t}} \tau_2 \\
 \coctx{\Gamma_1, \Gamma, \Gamma_2}{\cclrd{s}\,\cseq\,\cclrd{S}} \vdash \subst{e_2}{x}{e_s} : \tau_1
\end{array}\quad(*)
\]
Now using the (\emph{app}) rule and the fact that $\cpar\,=\,\cseq$, associativity, commutativity and idempotence
(note that all three properties are needed):
\[
\inference
 { (*) }
 {\inference
   { \coctx{\Gamma_1,\Gamma,\Gamma_2}{(\cclrd{r}\,\cseq\,\cclrd{S}) \;\cpar\; ((\cclrd{s}\,\cseq\,\cclrd{S}) \,\cseq\, \cclrd{t})} 
       \vdash \subst{e_1}{x}{e_s}~\subst{e_2}{x}{e_s} : \tau_2}
   { \coctx{\Gamma_1,\Gamma,\Gamma_2}{(\cclrd{r} \;\cpar\; (\cclrd{s} \,\cseq\, \cclrd{t}))\;\cseq\;\cclrd{S}} 
       \vdash \subst{(e_1~e_2)}{x}{e_s} : \tau_2} }
\]

\paragraph{(let)} Given the following typing derivation using (\emph{let}):
\[
\inference
  { \coctx{\Gamma_1,x:\tau_s,\Gamma_2}{\cclrd{r}} \vdash e_1 : \tau_1 &
    \coctx{\Gamma_1,x:\tau_s,\Gamma_2, y:\tau_1}{\cclrd{s}} \vdash e_2 : \tau_2}
  {\coctx{\Gamma_1,x:\tau_s,\Gamma_2}{\cclrd{s} \;\cpar\; (\cclrd{s} \,\cseq\, \cclrd{r})} \vdash \kvd{let}~y=e_1~\kvd{in}~e_2 : \tau_2 }
\]
From the induction hypothesis, we have that:
\[
\begin{array}{l}
 \coctx{\Gamma_1, \Gamma, \Gamma_2}{\cclrd{r}\,\cseq\,\cclrd{S}} \vdash \subst{e_1}{x}{e_s} : \tau_1 \\
 \coctx{\Gamma_1, \Gamma, \Gamma_2, y:\tau_1}{\cclrd{s}\,\cseq\,\cclrd{S}} \vdash \subst{e_2}{x}{e_s} : \tau_2
\end{array}\quad(\dagger)
\]
Now using the (\emph{let}) rule and similarly to the (\emph{app}) case:
\[
\hspace{-2em}
\inference
 { (\dagger) }
 { \inference
   { \coctx{\Gamma_1, \Gamma, \Gamma_2}{(\cclrd{s}\,\cseq\,\cclrd{S}) \;\cpar\; ((\cclrd{s}\,\cseq\,\cclrd{S}) \,\cseq\, (\cclrd{r}\,\cseq\,\cclrd{S}))} 
         \vdash \kvd{let}~y=\subst{e_1}{x}{e_s}~\kvd{in}~\subst{e_2}{x}{e_s} : \tau_2 } 
   { \coctx{\Gamma_1, \Gamma, \Gamma_2}{(\cclrd{s} \;\cpar\; (\cclrd{s} \,\cseq\, \cclrd{r})) \,\cseq\, \cclrd{S}} 
         \vdash \subst{(\kvd{let}~y=e_1~\kvd{in}~e_2)}{x}{e_s} : \tau_2 } }
\]
\end{proof}
\newpage

% =================================================================================================
%                                                                      
%     ###    #                           #                           ##   
%    #   #   #                           #                            #   
%    #      ####   # ##   #   #   ###   ####   #   #  # ##    ###     #   
%     ###    #     ##  #  #   #  #   #   #     #   #  ##  #      #    #   
%        #   #     #      #   #  #       #     #   #  #       ####    #   
%    #   #   #  #  #      #  ##  #   #   #  #  #  ##  #      #   #    #   
%     ###     ##   #       ## #   ###     ##    ## #  #       ####   ###  
%
% =================================================================================================

\section{Substitution for structural coeffects}
\label{sec:appendix-struct-cbn}

In order to prove that $\beta$-reduction and $\eta$-expansion preserve the type of an expression
in Section~\ref{sec:struct-syntactic-subst}, we required a multi-nary form of the substitution 
lemma (Lemma~\ref{thm:structural-substitution}). This section provides the corresponding proof.

\begin{lemma*}[Multi-nary substitution]
Given an expression with multiple holes filled by variables $x_{Ti}\!:\!\tau_{Ti}$ with coeffects $\cclrd{s_k}$:
%
\begin{equation*}
\coctx{\Gamma}{\aclrd{\textbf{r}}}~[\,\coctx{x_{T1}\!:\!\tau_{T1}}{\alift{\cclrd{s_1}}} \,|\, \ldots \,|\,
  \coctx{x_{Tk}\!:\!\tau_{Tk}}{\alift{\cclrd{s_k}}}\,] \vdash e_r : \tau_r
\end{equation*}
%
and a expressions $e_{Ti}$ with free-variable contexts $\Gamma_{Ti}$ annotated with $\aclrd{\mathbf{T_i}}$:
%
\begin{equation*}
\coctx{\Gamma_1}{\aclrd{\mathbf{T_1}}} \vdash e_{T1} : \tau_{T1}
\quad \ldots \quad
\coctx{\Gamma_k}{\aclrd{\mathbf{T_k}}} \vdash e_{Tk} : \tau_{Tk}
\end{equation*}
%
then substituting the expressions $e_{Ti}$ for variables $x_{Ti}$ results in an expression with a context
where the original holes are filled by contexts $\Gamma_{Ti}$ with coeffects $\cclrd{s_i} \aseq \aclrd{\mathbf{T_i}}$:
%
\begin{equation*}
\coctx{\Gamma}{\aclrd{\textbf{r}}}~[\,\coctx{\Gamma_{T1}}{\cclrd{s_1}\aseq\,\aclrd{\textbf{T}_1}} \,|\, \ldots \,|\, 
  \coctx{\Gamma_{Tk}}{\cclrd{s_k}\aseq\,\aclrd{\textbf{T}_k}}\,] \vdash \subst{\subst{e_r}{x_{T1}}{e_{T1}}\ldots}{x_{Tk}}{e_{Tk}} : \tau_r
\end{equation*}
\end{lemma*}
\begin{proof}
Assume that $\coctx{\Gamma}{\aclrd{\mathbf{T_i}}} \vdash e_{Ti} : \tau_{Ti}$ and we are substituting terms $e_{Ti}$ for 
variables $x_{Ti}$. Furthermore, we assume that are variables that are being substituted for actually appear in the original 
term (which means that $k$ is at most the number of variables). Note that we use upper-case $\aclrd{\textbf{T}}$ to distinguish 
the coeffect of the expression that is being substituted into an expression. Using structural induction over $\vdash$:

% -------------------------------------------------------------------------------------------------

\paragraph{Syntax-driven typing rules}
\paragraph{(var)} Given the following derivation using (\emph{var}):
\[
\inference
  { }
  {\coctx{y:\tau}{\alift{\cunit}} \vdash y : \tau }
\]
Here, the context contains exactly one variable and so $k=1$. There are two cases depending on whether $y$ is 
the (only substituted) variables $x_{T1}$ or not:
\begin{itemize}
\item If $y=x_{T1}$ then also $\tau = \tau_{T1}$ and thus $\subst{y}{x_{T1}}{e_{T_1}} = e_{T1}$. 
In this case, the context contains only a single hole $\coctx{\Gamma}{\aclrd{\mathbf{r}}} = \coctx{-}{-}$. Using the
assumption and the fact that $\cunit$ is a unit of $\cseq$:
\[
\inference
 { \inference 
   {\coctx{\Gamma_{T1}}{\aclrd{\mathbf{T_i}}} \vdash \subst{y}{x_{T1}}{e_{T1}} : \tau_r }
   {\coctx{\Gamma}{\aclrd{\mathbf{r}}}~[\,\coctx{\Gamma_{T1}}{\aclrd{\mathbf{T_i}}}\,] \vdash \subst{y}{x_{T1}}{e_{T1}} : \tau_r } }
 { \coctx{\Gamma}{\aclrd{\mathbf{r}}}~[\,\coctx{\Gamma_{T1}}{\cunit\,\aseq\,\aclrd{\mathbf{T_i}}}\,] \vdash \subst{y}{x_{T1}}{e_{T1}} : \tau_r }   
\]
\item If $y\neq x_{T1}$ then there is no substitution that could be performed ($y$ does not 
appear in the context) and so (trivially):
\[
\coctx{\Gamma}{\aclrd{\mathbf{r}}}~[\,] \vdash y : \tau_r
\]
\end{itemize}

\paragraph{(app)} In the (\emph{app}) rule, the context $\Gamma$ is obtained as a tensor product $\Gamma_1, \Gamma_2$.
  Given $\Gamma$ of length $k$, we assume that $\Gamma_1$ has length $l$ and $\Gamma_2$ has length $k-l$. 
  Now, given the following typing derivation using (\emph{app}):
\[
\inference
  {\coctx{\Gamma_1}{\aclrd{\mathbf{s_1}}} 
   [\,\coctx{x_{T1}}{\alift{\cclrd{s_1}}} \,|\, \ldots \,|\, \coctx{x_{Tl}}{\alift{\cclrd{s_l}}}\,] 
      \vdash e_1 : \tau_1 \xrightarrow{\cclrd{t}} \tau_2 \\
   \coctx{\Gamma_2}{\aclrd{\mathbf{s_2}}} 
   [\,\coctx{x_{Tl+1}}{\alift{\cclrd{s_{l+1}}}} \,|\, \ldots \,|\, \coctx{x_{Tk}}{\alift{\cclrd{s_k}}}\,]
      \vdash e_2 : \tau_1 }
  { \begin{array}{l}
    \coctx{\Gamma_1, \Gamma_2}{\aclrd{\mathbf{s_1}\;\atimes\;(\cclrd{t} \,\aseq\, \aclrd{\mathbf{s_2}})}}~
     [\,\coctx{x_{T1}}{\alift{\cclrd{s_1}}} \,|\, \ldots \,|\, \coctx{x_{Tl}}{\alift{\cclrd{s_l}}}, \\[-0.35em] \hspace{9.3em}
       \coctx{x_{Tl+1}}{\alift{\cclrd{t} \,\cseq\, \cclrd{s_{l+1}}}} \,|\, \ldots \,|\, 
          \coctx{x_{Tk}}{\alift{\cclrd{t} \,\cseq\, \cclrd{s_k}}}\,] \vdash e_1~e_2 : \tau_2 
    \end{array} }
\]
Here, we use $\cclrd{t}\;\aseq\;\aclrd{\mathbf{s_2}}$ as a pointwise extension of $\czip$ that is additionally
applied to holes such that $\cclrd{t}\;\czip\; - = -$. That is, a hole in the context remains a hole and so
the second part of the context is filled with $\cclrd{t} \,\cseq\, \cclrd{s_i}$ for all $i \in \{ l+1 \ldots k \}$.
Next, from the induction hypothesis, we have that:
\[
\begin{array}{l}
 \coctx{\Gamma_1}{\aclrd{\mathbf{s_1}}} 
   [\,\coctx{\Gamma_{T1}}{\cclrd{s_1}\,\aseq\,\aclrd{\mathbf{T_1}}} \,|\, \ldots \,|\, \coctx{\Gamma_{Tl}}{\cclrd{s_l}\aseq\,\aclrd{\mathbf{T_l}}}\,] 
   \vdash e_1~[\,\ldots\,] : \tau_1 \xrightarrow{\cclrd{t}} \tau_2 \\
 \coctx{\Gamma_2}{\aclrd{\mathbf{s_{2}}}} 
   [\,\coctx{\Gamma_{Tl+1}}{\cclrd{s_{l+1}}\aseq\,\aclrd{\mathbf{T}_{l+1}}} \,|\, \ldots \,|\, \coctx{\Gamma_{Tk}}{\cclrd{s_k}\aseq\,\aclrd{\mathbf{T_k}}}\,]
   \vdash e_2~[\,\ldots\,] : \tau_1
\end{array}\quad(*)
\]
Now using the (\emph{app}) rule in the first step and associativity of $\cseq$ on the second part of the context in the second step:
\[
\hspace{-2em}
\inference 
 { \inference
   { (*) }
   { \begin{array}{l}
      \coctx{\Gamma_1, \Gamma_2}{\aclrd{\mathbf{s_1}\;\atimes\;(\cclrd{t} \,\aseq\, \aclrd{\mathbf{s_2}})}} ~~
       [\,\coctx{\Gamma_{T1}}{ \cclrd{s_1}\,\aseq\,\aclrd{\mathbf{T_1}} } \,|\, \ldots \,|\, 
          \coctx{\Gamma_{Tl}}{ \cclrd{s_l}\,\aseq\,\aclrd{\mathbf{T_l}} }, 
         \\[-0.35em] \qquad
         \coctx{\Gamma_{Tl+1}}{\cclrd{t} \,\aseq\, (\cclrd{s_{l+1}}\,\aseq\,\aclrd{\mathbf{T_{l+1}}}) } \,|\, \ldots \,|\,
         \coctx{\Gamma_{Tk}}{\cclrd{t} \,\aseq\, (\cclrd{s_k}\,\aseq\,\aclrd{\mathbf{T_k}})}\,] \vdash (e_1~e_2)~[\,\ldots\,] : \tau_2 
      \end{array} } }
 { \begin{array}{l}
    \coctx{\Gamma_1, \Gamma_2}{\aclrd{\mathbf{s_1}\;\atimes\;(\cclrd{t} \,\aseq\, \aclrd{\mathbf{s_2}})}} ~~
     [\,\coctx{\Gamma_{T1}}{ \cclrd{s_1}\,\aseq\,\aclrd{\mathbf{T_1}} } \,|\, \ldots \,|\, 
        \coctx{\Gamma_{Tl}}{ \cclrd{s_l}\,\aseq\,\aclrd{\mathbf{T_l}} }, 
       \\[-0.35em] \qquad
       \coctx{\Gamma_{Tl+1}}{(\cclrd{t} \,\cseq\, \cclrd{s_{l+1}})\,\aseq\,\aclrd{\mathbf{T_{l+1}}} } \,|\, \ldots \,|\,
       \coctx{\Gamma_{Tk}}{(\cclrd{t} \,\cseq\, \cclrd{s_k})\,\aseq\,\aclrd{\mathbf{T_k}}}\,] \vdash (e_1~e_2)~[\,\ldots\,] : \tau_2 
    \end{array} }
\]

\paragraph{(abs)} Without the loss of generality, we can assume that the bound variable is 
not one of the variables being substituted. Thus, the last variable (and the corresponding coeffect)
are not holes. The typing derivation using (\emph{abs}) then looks as follows:
\[
\inference
  {\coctx{\Gamma, x\!:\!\tau_1}{\aclrd{\textbf{r}}\,\atimes\,\alift{\cclrd{s}}} 
   ~[\,\coctx{x_{T1}\!:\!\tau_{T1}}{\alift{\cclrd{s_1}}} \,|\, \ldots \,|\, \coctx{x_{Tk}\!:\!\tau_{Tk}}{\alift{\cclrd{s_k}}}\,]
    \vdash e : \tau_2}
  {\coctx{\Gamma}{\aclrd{\textbf{r}}} 
   ~[\,\coctx{x_{T1}\!:\!\tau_{T1}}{\alift{\cclrd{s_1}}} \,|\, \ldots \,|\, \coctx{x_{Tk}\!:\!\tau_{Tk}}{\alift{\cclrd{s_k}}}\,]
    \vdash \lambda x.e : \tau_1 \xrightarrow{\cclrd{s}} \tau_2 }
\]
From the induction hypothesis, we have that:
\[
\coctx{\Gamma, x\!:\!\tau_1}{\aclrd{\textbf{r}}\,\atimes\,\alift{\cclrd{s}}}~
 [\,\coctx{\Gamma_{T1}}{\cclrd{s_1}\aseq\,\aclrd{\textbf{T}_1}} \,|\, \ldots \,|\, 
    \coctx{\Gamma_{Tk}}{\cclrd{s_k}\aseq\,\aclrd{\textbf{T}_k}}\,] \vdash e~[\,\ldots\,] : \tau_2
\]
Because the last position in the vector of variables is an actual variable rather than a hole, we
just need to apply the (\emph{abs}) rule:
\[
\inference 
 { \coctx{\Gamma, x\!:\!\tau_1}{\aclrd{\textbf{r}}\,\atimes\,\alift{\cclrd{s}}}~
   [\,\coctx{\Gamma_{T1}}{\cclrd{s_1}\aseq\,\aclrd{\textbf{T}_1}} \,|\, \ldots \,|\, 
      \coctx{\Gamma_{Tk}}{\cclrd{s_k}\aseq\,\aclrd{\textbf{T}_k}}\,] \vdash e~[\,\ldots\,] : \tau_2 }
 { \coctx{\Gamma}{\aclrd{\textbf{r}}}~
   [\,\coctx{\Gamma_{T1}}{\cclrd{s_1}\aseq\,\aclrd{\textbf{T}_1}} \,|\, \ldots \,|\, 
      \coctx{\Gamma_{Tk}}{\cclrd{s_k}\aseq\,\aclrd{\textbf{T}_k}}\,] \vdash \lambda x.e~[\,\ldots\,] : \tau_1 \xrightarrow{\cclrd{s}} \tau_2 }
\]

\paragraph{(let)} In the structural coeffect calculus let-binding can be viewed as a syntactic
sugar for abstraction/application and so the case follows from (\emph{abs}) and (\emph{app}).

\vspace{0.75em}
\noindent
Compared to the similar proof for the flat coeffect calculus, the proof for the structural system
requires fewer properties of the coeffect algebra. In particular, we only needed associativity of 
$\cseq$ in the (\emph{app}) rule the fact that $\cunit$ is a unit of $\cseq$ in (\emph{var}).

% -------------------------------------------------------------------------------------------------

\paragraph{Structural rules}

\paragraph{(contr)} In case of contraction, we can assume that the two variables to be contracted
(in the assumption) are not the variables that are being substituted. However, the resulting
variable  (in the conclusion) can be one of the variables being substituted for.
